The reference model approach in feature selection problems \vspace{.1in}

### 2019-07-16 paper/ref_approach.tex
Update ref_approach.tex
rewrite the introduction

@@

In statistical applications, one of the main steps of the modelling
workflow is covariate or feature selection, which is a special case of
model reduction.  Simpler models bring advantages in reduced
measurement costs and improved interpretability. Variable selection
can be also used as an inferential tool: the interest is on knowing
which variables contain predictive information about the target
variable of interest.  The variable selection is most challenging in a
\textit{small n large p} scenario, which is common case, for example,
in microarray data analysis in medicine and biology.

@@

The goal of the variable selection can be the minimal subset which
provides good predictive performance (or balance sparsity with
predictive performance), or the entire collection of variables which have some
predictive information about the target variable.  In the predictive
performance based variable selection, the future data generation
mechanism can be modelled either using the observed data or a
reference model.
% process is going to be based on assumptions regarding the ``true''
% data generation mechanism.  Making assumptions means making models,
% thus we identify two families of feature selection, or more in general
% model reduction, approaches depending on how the data generation
% mechanism is modelled: the data based and the reference model based
% ones.
In the data based approaches, such as the Lasso selection
\citep{tibshirani1996regression} and the stepwise backward/forward
regression, the observed empirical data distribution is used as a
proxy of the future data usually in combination with cross-validation
or information criteria to take into account the double use of the
data.
%
The reference model based methods approximate the future data generation
mechanism using the predictive distribution of a reference model,
which can be, for example, the full-encompassing model. The reference
model approach has been used for long time in the Bayesian framework
since seminal work by \citet{paper:reference_lindley}. For more
historical references see the review by \citet{vehtari2012survey}, and for
most recent methodological development see 
\citet{paper:projpred}.

@@

The reference models have been used also in non-Bayesian context. For example,
\cite{harrell2015regression} refers to the reference model as a full model that can be thought as the ``gold standard''.
% and shows some examples of how it can be used to seek a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model.
\cite{faraggi2001understanding} deal with the necessity of identifying
interpretable risk groups in the context of survival data using neural
networks, which typically perform very well in terms of prediction,
but whose covariates are difficult to be understood in terms of
relevance. % \cite{faraggi2001understanding} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups.
\cite{paul2008preconditioning}, using the term preconditioning,
explore approximating models fitting Lasso or stepwise regression
against consistent estimates $\hat{y}$ of a reference model instead of
the observed response quantities $y$.
%\cite{harrell2015regression} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
\\
%\cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance of the explored submodels with the reference model one.

%\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and, when it is used, on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

In this paper we examine in detail the benefits of using a reference
model.  A properly designed reference model is able to filter part of
the noise present in the data, hence to provide an improved and more
stable selection process even if the reference model would not be the
true data generating process. The substantial reduction of the variance is more important than small bias due the potential model misspecification.

We argue that, however the reference model is used, it can be always seen acting as a filter on the observed data. Following this idea, we devise a simple reference model approach that can be applied on top of any selection procedure. We thus show the reference model benefits, regardless of what specific procedure is applied, using as a benchmark the normal means estimation problem. We apply different state-of-the-art methods comparing in each case the results basing the selection on the data or on the reference model. In case of full Bayesian methods, we report results in terms of shrinkage and accuracy of the posterior estimates, whereas when a full selection is carried out, we compute quantities as false discovery rate, sensitivity and a measure of the stability of the selection. We also demonstrate the improvements in applied case when using the reference model based projection predictive approach versus the stepwise regression.

The results indicate the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognise that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures.
% We hope that this work can increase
% the attractiveness on this family of methods and, therefore, it can be
% of interest to a large number of practitioners.
\\
We would like also the reader to note that although we focus on
Bayesian approach, the benefits of the reference model approach can be
obtained also in frequentist settings.

The paper is structured as follows. In Section \ref{reference-model-approach} we review the concept of the reference model, its benefits with examples, including the projection predictive approach, and how it can be used as a filter on data in a simple way. In Section \ref{comparison} we use different methods for feature selection comparing the selection results using or not the reference model in the framework of the normal means problem. Finally, in Section \ref{conclusion} we present our conclusions.

>>>>

In statistical applications, one of the main steps in the modeling
workflow is covariate or feature selection, which is a special case of
model reduction. Variable selection may have multiple goals. 
First, if the covariates themselves are of interest,
we can use variable selection as an inferential tool to find the entire 
collection of variables which have some predictive information about the target variable.
Second, as simpler models come with the advantages of reduced
measurement costs and improved interpretability, we may be interested in
finding the minimal subset of covariates which
provides good predictive performance (or good balance between simplicity and
predictive performance). 
In the latter case, where the focus is on prediction, the true data generation
mechanism of future data can be approximated either by using the observed data directly or
alternatively by using predictions from a \emph{reference model}.

In data based approaches, such as Lasso selection
\citep{tibshirani1996regression} or stepwise backward/forward
regression, the observed empirical data distribution is utilized as a
proxy of future data usually in combination with cross-validation
or information criteria to provide estimates for out-of-sample
predictive performance.
In contrast, reference model based methods approximate the future data generation
mechanism using the predictive distribution of a reference model,
which can be, for example, a full-encompassing model including all
covariates. The reference
model approach has been used for long time in Bayesian statistics
since the seminal work of \citet{paper:reference_lindley}. For more
historical references, see the review by \citet{vehtari2012survey}, and for
most recent methodological development see 
\citet{paper:projpred}.
Reference models have been also used in non-Bayesian contexts,
where \cite{harrell2015regression} describes them as full models that
can be thought of as a ``gold standard'' (for a given application).
For example,
\cite{faraggi2001understanding} deal with the necessity of identifying
interpretable risk groups in the context of survival data using neural
networks, which typically perform very well in terms of prediction,
but whose covariates are difficult to be understood in terms of
relevance. 
\cite{paul2008preconditioning}, using the term preconditioning,
explore approximating models fitting Lasso or stepwise regression
against consistent estimates $\hat{y}$ of a reference model instead of
the observed responses $y$.

In this paper, we examine the benefits of using reference
models in detail. We illustrate how a properly designed reference model is able 
to filter parts of the noise present in the data, and hence provides an improved and more
stable selection process. This holds even if the reference model does not actually resemble the
true data generating process. Moreover, our analyses indicate that the substantial reduction of
variance attributable to noise is usually
more important than small potential bias due model misspecification.
We argue that, regardless of how the reference model is set up and used in the
inference prodecure, it can be always seen as acting as a filter on the observed 
data. Following this idea, we devise a simple reference model approach that can
be applied on top of any variable selection procedure.

In our simulations and analysis of real data, we apply several state-of-the-art methods for variable selection and compare,
for each of these methods, the results obtained by using a data based and a
reference model based approach. When a full 
variable selection is carried out, we report and compare quantities such as 
the false discovery rate, sensitivity and stability of the selection. For fully Bayesian methods, we additionally report results in terms of shrinkage and accuracy of the posterior estimates.
We show the benefits
of a reference model approach, regardless of what specific procedure is applied, using 
the normal means estimation problem as a benchmark. 
Variable selection is typically most challenging in 
\textit{small n large p} scenarios, that is, scenarios with small number of
observations but a lot of unknown model parameters. These occur frequently, for example, in microarray data analysis in medicine or biology.
In general, the less data we have and the more complex 
the estimated models, the higher the benefit of using a reference
models as the basis for variable selection.
Our results indicate that the core reason why the reference model based
methods perform well is the reference model itself, rather than the
specific way of using it. We do recognise that all these methods
diversify how the subset of variables is actually chosen and which
properties it has, thus it remains fundamental to investigate and
develop such different procedures. We want to note that, although we focus on
Bayesian approach in this paper, the benefits of the reference model 
approach can also be obtained in non-Bayesian settings.

The remainder of the paper is structured as follows. In Section \ref{reference-model-approach}, we review the concept of the reference model, its benefits with examples, including the projection predictive approach, and how it can be used as a filter on data in a simple way. In Section \ref{comparison}, we use different methods for feature selection comparing the selection results with and without using a reference model in the framework of the normal means problem. Finally, we end with a conclusion in Section \ref{conclusion}.

@@

@@

% process is going to be based on assumptions regarding the ``true''
% data generation mechanism.  Making assumptions means making models,
% thus we identify two families of feature selection, or more in general
% model reduction, approaches depending on how the data generation
% mechanism is modelled: the data based and the reference model based
% ones.

@@

% and shows some examples of how it can be used to seek a sparser approximation. He highlights some of the benefits as the possibility to calculate the accuracy with which the submodel approximates the best model and the inheritance of the shrinkage when it is applied to the full model.

% \cite{faraggi2001understanding} propose to combine the benefits of the neural networks, which become what we call reference models, with the interpretability of regression trees, in order to provide a framework to make inferences about risk groups.

%\cite{harrell2015regression} includes also the possibility of using simple least squares for the approximating models as long as the target is the linear predictor of a full regression model \hl{CHECK}.
%\cite{paper:projpred} extend the projection algorithm with the clustered projection and choose the best model comparing the predictive performance of the explored submodels with the reference model one.
%\cite{paper:projpred} show in their experiments the benefits of the projection predictive approach in terms of predictive ability of the selected model and smaller size compared to other methods. It could be of interest to know how often a certain method includes non-relevant features. One could argue that such question can not have a general answer, because it depends both on the data and, when it is used, on the choice of the reference model. Moreover in real settings it is hard to define what relevant means, in most of the cases all the features have some level of relevance in term of prediction ability and thus the question looses of meaning.  However, it is still possible to simulate certain artificial scenarios and control which features are included. \cite{paper:model_selection} report results about the proportion of relevant and non-relevant features selected with different variable selection methods, showing overall better results in favour of the projection predictive approach. The explored submodels using the projection show also a higher performance and stability with respect to the compared methods, given the same submodel size.

% We hope that this work can increase
% the attractiveness on this family of methods and, therefore, it can be
% of interest to a large number of practitioners.
